---
title: "1-tidyverse"
author: "Dan Shah"
date: "1/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# A VERY fast introduction to the Tidyverse

### If you haven't used the Tidyverse before I recommend:

-   Ch 1-4 of <https://moderndive.netlify.app/>or

-   <https://r4ds.had.co.nz> for a deeper dive

Let's load the required packages and see what data is included:

```{r load, echo = FALSE, warning= FALSE}
#load required packages
library(nycflights13)
library(tidyverse)
library(lubridate)

data(package='nycflights13')
```

### Examining the table

You can also embed plots, for example:

```{r flightsView, echo=FALSE}
#View(flights)
```

```{r flightsGlimpse, echo=FALSE}
glimpse(flights)
```

```{r help}
?flights
```

### Filter

Flights has more than one carrier (airline) within the table.

```{r allCarrier}
unique(flights$carrier)
```

To filter for one (or more) the filter function can be used.

Filter for flights from United Airlines (UA):

```{r filter}
ua_flights<- flights %>% 

  filter(carrier == "UA")

unique(ua_flights$carrier)
```

### Select and Mutate

```{r select}
flights %>% 
  select(year, month, day, dep_time, flight, carrier) %>% 
  head()
```

Mutate creates new columns within the dataframe. The make_date function from the lubridate package makes creating date or datetime vectors within a dataframe quite easy. Here we use the integer columns from year, month day to make a new column called date_flight:

```{r mutate}
flights %>% 
  select(year, month, day) %>% 
  mutate(date_flight = make_date(year, month, day)) %>% 
  head()
```

### **Group By, Summarise and Arrange**

Group By is a powerful function from the dplyr package within the tidyverse that groups data frames based upon keys:

```{r groupby}
flights_grouped_carrier <- flights %>% 
  group_by(carrier)

flights_grouped_carrier
```

From there we can obtain specifics within the group, like number of rows, average dep_delay:

```{r summarise}
flights_grouped_carrier %>% 
  summarise(cnt = n(),
            avg_delay = mean(dep_delay, na.rm = T)) %>% 
  arrange(desc(avg_delay))
```

### Visualization with ggplot2

<https://r4ds.had.co.nz/data-visualisation.html>

1.  Begin a flight with the ggplot()

2.  Include the dataset

3.  Choose how to map the data using the aes() aesthetics function

4.  Add layers, like a geom layer, which generates the plot and has additional settings

`ggplot(data = <DATA>) +`

`<GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))`

If we wanted to create a scatter plot with the ua_flights dataset and dep_delay on the x, arr_delay on the y, then we would use:

```{r scatterUA, warning = FALSE}
ggplot(data = ua_flights, mapping = aes(x = dep_delay, y = arr_delay)) + 

  geom_point(alpha = 0.2)
```

### Conditional filtering

Use \| and & to conditionally filter a dataset:

```{r conditionalFilter}
early_january_weather <- weather %>% 

  filter(origin == "LGA" & (month == 1 & day <= 15))

head(early_january_weather)
```

### Plotting a time series:

```{r timeseriesPlot}
ggplot(data = early_january_weather, 
       mapping = aes(x = time_hour, y = temp)) + 
  geom_line() +
  ggtitle((title = "Weather in early January")) +
  ylab("Temperature (F)") +
  xlab("Datetime")
```

Next lecture we'll cover tsibble objects (we're working with tibble objects today) and the autoplot() function which simplifies things:

```{r autoplot}
library(tsibble)
library(feasts)

early_january_weather %>% 
  tsibble(index = time_hour) %>% 
  autoplot(temp)
```

### Lab 1

1.  What kind of object is early_january_weather (use str())?

2.  Plot wind speeds (weather table) in February by origin? What stands out?

3.  What month, on average, had the highest wind speed with origin = JFK
