---
title: "12_Prophet"
author: "Dan Shah"
date: "4/13/2022"
output: html_document
---

## Project Tips

Importing a csv (or text file) from the web by pasting the url as a string to read.csv()

```{r csv}
url <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/births/US_births_2000-2014_SSA.csv"
df_csv_file <- read.csv(url, header = TRUE) #if .tsv file read.delim(url, sep = "/t", header = TRUE)
print(df_csv_file)
```

## FRED

For FRED data you can copy and paste the link to the csv data as a url and follow the same process

```{r fred}
url <- "https://fred.stlouisfed.org/graph/fredgraph.csv?bgcolor=%23e1e9f0&chart_type=line&drp=0&fo=open%20sans&graph_bgcolor=%23ffffff&height=450&mode=fred&recession_bars=on&txtcolor=%23444444&ts=12&tts=12&width=1168&nt=0&thu=0&trc=0&show_legend=yes&show_axis_titles=yes&show_tooltip=yes&id=DFF&scale=left&cosd=2017-04-12&coed=2022-04-12&line_color=%234572a7&link_values=false&line_style=solid&mark_type=none&mw=3&lw=2&ost=-99999&oet=99999&mma=0&fml=a&fq=Daily%2C%207-Day&fam=avg&fgst=lin&fgsnd=2020-02-01&line_index=1&transformation=lin&vintage_date=2022-04-13&revision_date=2022-04-13&nd=1954-07-01"

fred_csv_file <- read.csv(url, header = TRUE)
print(fred_csv_file)
```
You can also use the fredr package for FRED data which allows you to interact
with the application (API) for FRED:

```{r}
#install.packages('fredr')
library(fredr)
fredr_set_key('YOU NEED TO CREATE AN ACCOUNT AND GET AN API KEY')

library(dplyr)
library(ggplot2)

popular_funds_series <- fredr_series_search_text(
    search_text = "federal funds",
    order_by = "popularity",
    sort_order = "desc",
    limit = 1
)

popular_funds_series_id <- popular_funds_series$id

popular_funds_series_id %>%
  fredr(
    observation_start = as.Date("1990-01-01"),
    observation_end = as.Date("2000-01-01")
  ) %>%
  ggplot(data = ., mapping = aes(x = date, y = value, color = series_id)) +
    geom_line() +
    labs(x = "Observation Date", y = "Rate", color = "Series")

```

## Stock data

Why I'm advising against forecasting only stock data:
https://towardsdatascience.com/quit-trying-to-predict-the-market-27d77149a709

It's more or less random noise once made stationary:

```{r}
library(fpp3)
google_2015 <- gafa_stock %>%
  filter(Symbol == "GOOG", year(Date) == 2015) 
google_2015 %>% ACF(Close) %>% 
  autoplot() + labs(subtitle = "Google closing stock price")
google_2015 %>% ACF(difference(Close)) %>% 
  autoplot() + labs(subtitle = "Changes in Google closing stock price")
```

## Number of Time Series

The # of time series within a dataset was answered incorrectly for most submissions.
I just want to clarify that many of these datasets have many time series. The below has 304 time series data:

```{r}
tourism
```
But feasts and fable packages handle these time series without the need for writing loops.

Here's how you can explore these time series data quickly: 
trend/seasonality for all time series (all 304) within this dataset

```{r}
system.time(
  tourism_features <- tourism %>% 
    features(Trips, feature_set(pkgs = "feasts")))

tourism_features
  

```

The packages can handle modeling many time series at once:

```{r}
system.time(
  fit <- tourism %>% 
    model(ETS(Trips))
  
)
```


## Prophet

The package that integrates with fable can be installed and loaded:

```{r}
#install.packages("fable.prophet")
library(fable.prophet)
```

Let's import and clean wikipedia visit data for Peyton Manning

```{r}
url <- 'https://raw.githubusercontent.com/facebook/prophet/main/examples/example_wp_log_peyton_manning.csv'
pm <- read.csv(url, header = TRUE)
pm$Date <- ymd(pm$ds)
pm <- tsibble(pm, index = Date)
pm$Visits <- pm$y

#missing dates
pm <- pm %>% fill_gaps() %>% tidyr::fill(Visits, .direction = "down")


```

Let's visualize the data:
```{r}
pm %>% autoplot()
```
We can compare accuracy to ETS and ARIMA:

```{r}
#split into train/test
train <- pm %>% filter(year(Date) < 2015)
test <- pm %>% filter(year(Date) >= 2015)
```


```{r}
fit <- train %>% 
  model(prophet(Visits),
        ARIMA(Visits),
        ETS(Visits))
```



```{r}
fc <- fit %>% forecast(h = 385)
fc %>% accuracy(test)
```

Decompose prophet model
```{r}
fit_prophet <- train %>% 
  model(prophet(Visits))

fit_prophet %>% components() %>% autoplot()
```
We can also use the base Prophet package, but it doesn't play well with Fable:

```{r}
#install.packages('prophet')
library(prophet)

```

Let's reload the data and plot the forecast vs the actuals:
```{r}
url <- 'https://raw.githubusercontent.com/facebook/prophet/main/examples/example_wp_log_peyton_manning.csv'
pm <- read.csv(url, header = TRUE)
fit_prophet <- prophet(pm)
future <- make_future_dataframe(fit_prophet, periods = 365)
fc <- predict(fit_prophet, future)
plot(fit_prophet, fc)
```
Can plot components with the base package also:

```{r}
prophet_plot_components(fit_prophet, fc)
```

