---
title: "2-visualization"
author: "Dan Shah"
date: "1/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(fpp3)

```

## tsibble

tsibble objects

```{r}
olympic_running
```

Creating a tsibble from scratch

```{r}
t <- tsibble(
  Year = 2001:2022,
  Value = rnorm(n = 22),
  index = Year
)

str(t)
```

```{r}
autoplot(t, Value)
```

Reading in a file and converting to a tsibble

```{r}
prison <- readr::read_csv("https://OTexts.com/fpp3/extrafiles/prison_population.csv")
prison
```

```{r}
#reading in a file and converting to tsibble

prison <- prison %>%
  mutate(Quarter = yearquarter(Date)) %>%
  select(-Date) %>%
  as_tsibble(key = c(State, Gender, Legal, Indigenous),
             index = Quarter)

prison
```

```{r}
index(prison)
```

```{r}
key(prison)
```

The dplyr functions work with tsibbles:

```{r}
prison %>% 
  filter(State == "NT")
```

### **Visualizing Time Series Data**

autoplot

```{r}
#plotting a time series
melsyd_economy <- ansett %>%
  filter(Airports == "MEL-SYD", Class == "Economy") %>%
  mutate(Passengers = Passengers/1000)

autoplot(melsyd_economy, Passengers) +
  labs(title = "Ansett airlines economy class",
       subtitle = "Melbourne-Sydney",
       y = "Passengers ('000)")
```

Information regarding autoplot

```{r}
?autoplot
```

Seasonal plots

```{r}
glimpse(vic_elec)
```

```{r}
vic_elec_2012 <- vic_elec %>% filter(Time < as.Date('2013-01-01'))
vic_elec_2012 %>% autoplot(Demand)
```

```{r}
vic_elec_2012 %>% gg_season(Demand, period = 'day')
```

```{r}
vic_elec %>% gg_season(Demand, period = "week")
```

**Correlation and scatterplot matrices**

Beer and Tobacco

```{r}
aus_production
```

```{r}
aus_production %>% 
  ggplot(aes(x = Beer, y = Tobacco)) +
  geom_point() +
  labs(title = "Beer vs Tobacco")
```

```{r}
not_missing <- aus_production %>% 
  filter(!is.na(Tobacco))

cor(not_missing$Beer, not_missing$Tobacco)
```

**Lag and Autocorrelation plots**

```{r}
aus_production %>% 
  autoplot(Beer)
```

```{r}
#example from fpp3 2.7
recent_production <- aus_production %>% 
  filter(year(Quarter) >= 2000)

recent_production %>% 
  gg_lag(Beer, geom = "point") +
  labs(x = "lag(Beer, k")
```

```{r}
aus_production %>% 
  ACF(Beer, lag_max = 5)
```

```{r}
aus_production %>% 
  ACF(Beer) %>% 
  autoplot() 
```

```{r}
aus_production %>% 
  filter(year(Quarter) >= 2000) %>% 
  ACF(Beer) %>% 
  autoplot()
```

**White noise**

```{r}
set.seed(44)
noise <- tsibble(x = 1:100, white_noise = rnorm(100), index = x)
noise %>%
  autoplot(white_noise) +
  labs(title = "Random noise")
```

```{r}
noise %>% mutate(lag_noise = lag(white_noise), lag2_noise = lag(white_noise, 2))
```

```{r}
noise %>% ACF(white_noise)
```

```{r}
noise %>% 
  ACF(white_noise) %>% 
  autoplot() +
  labs(title = "Correlogram of White Noise")
```

**lubridate package**

```{r}
library(lubridate)
today()
class(today())
```

```{r}
now()
class(now())
```

```{r}
string_date <- "01/19/2022 7:00:02"
class(string_date)

datefrmt_date <- mdy_hms(string_date)
class(datefrmt_date)
```

### **Datetime components**

```{r}
current_dt <- now()
year(current_dt)
month(current_dt)
wday(current_dt)
day(current_dt)
wday(current_dt, label = TRUE, abbr = FALSE)
current_dt
```

```{r}
old_dt <- update(current_dt, year = 2020)
old_dt
```

Subtraction will return a

```{r}
difference <- current_dt-old_dt
difference
class(difference)
```

Could be different units so best to stick to a duration class

```{r}
as.duration(difference)
```

To create durations you can use functions:

```{r}
ddays(1) + current_dt
```

**Durations** are exact times. Sometimes it's best to use periods when things like daylight savings time is involved or a year is a leap year. Duration will not give expected results.

```{r}
days(1) + current_dt
```

**Timezones**

```{r}
Sys.timezone()

head(OlsonNames())
```

To update the timezone on a datetime use with_tz()

```{r}
with_tz(current_dt, "US/Pacific") 
#force_tz(current_dt, "US/Pacific") #will change the underlying data so only if tz is wrong
```

### Time Series manipulations

**upsampling**

```{r}
tourism
```

```{r}
index(tourism)
lubridate::year(yearquarter("2000 Q1"))
```

```{r}
tourism %>%
  mutate(Year = lubridate::year(Quarter)) %>%
  index_by(Year) %>%
  group_by(Region, State, Purpose) %>%
  summarise(Trips = sum(Trips)) %>%
  ungroup()
```

**shifting**

```{r}
tourism %>% 
  group_by(Region, State, Purpose) %>% 
  mutate(Trips_lag1 = lag(Trips)) %>% 
  ungroup()
```

**rolling with slider package**

```{r}
aus_production
```

```{r}
library(slider)
aus_production %>% 
  mutate(ma_beer = slide_dbl(Beer, mean, .before = 5))
```

**missing data, forward fill**

```{r}
tail(aus_production, 40)
```

```{r}
aus_production %>%
  tidyr::fill(Tobacco, .direction = 'down') %>% 
  tail()
```

### **Lab**

1.  What are the groupings and base time index for the olympic_running dataset?
2.  use autoplot() to view bricks production in aus_production dataset. Include labels and a title.
3.  Plot out bricks production with quarter on the x-axis and individual lines for the years in the dataset
4.  Produce a scatterplot of bricks and cement from aus_production to see the relationship
5.  Calculate correlation between Bricks and Cement columns Hint: must remove the missing values with !is.na() (try Googling if you're stumped)
6.  Plot the ACF for Bricks using the ACF() function
7.  What do the blue dashed lines on the plot above mean?
8.  What can you interpret from the plot?
